Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid, Xinhua and Diplomatic Visits

### 2022-09-29 04_paper/0929_draft.Rmd
adding aid section

Although Beijing has primarily focused on its immediate surrounding neighbors, Asia-Pacific and Central Asia, for security and geostrategic concerns, it is nevertheless increasingly thinking and acting globally [@nathan_chinas_2012]. Similarly, in terms of where China’s money goes and its economic influence is concentrated, some countries clearly are more important than others. However, Beijing's leaders hold the reasons why close to the vest.

@@

However, the goals of facilitating cooperation and winning international recognition as a rising power are difficult to accomplish in secret. This, therefore, raises a natural question. *Do China’s foreign aid and debt allocation relate to public diplomacy and international recognition efforts? If so, how?*

@@

As we test these hypotheses, in our research design we take care to ensure that we delineate China's development financing according to the type of assistance China offers. Until recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) have unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.^[See @dreher_banking_2022.] Thus, as we test our hypotheses, we also distinguish between Beijing-as-donor from Beijing-as-lender.

@@

Aid recipient countries disproportionately get more media attention from *Xinhua* via stories that report both on the state of their economy and their development needs. Notably, *Xinhua* does not just promote China’s efforts to offer aid to those countries in need but also gives a rounded picture of those countries’ economic conditions justifying their legitimate need for aid.

Meanwhile, the finding that loans correspond with diplomatic missions has special relevance in light of Beijing's Belt and Road Initiative (BRI). A 2018 BRI foreign policy assessment report from Washington [@hurley_examining_2019] highlights that lending indicates Beijing’s economic interests and strategic dominance---China tends to lend to vulnerable states with debt distress in exchange for unfair deals and strategic concessions. Brautigam [-@brautigam_critical_2020] calls the Chinese "debt-trap diplomacy" a meme based on collective fear of the Chinese threat rather than sound empirical evidence. Our finding suggests that China is more likely to provide loans to diplomatically aligned countries. For instance, Russia, one of China’s closest allies, receives the highest amount of lending on average among 142 countries but no aid at all between 2002 and 2017. In contrast, Iraq shows the opposite pattern: it receives the highest amount of aid from China during the period of our investigation, but it does not receive any loans. Concern about Chinese debt nonetheless raises a natural question related to the function of diplomacy in debtor countries. Does debt serve as a tool of diplomacy, or does Beijing wield diplomacy to cover a multitude of "debt-trap" sins? More research is needed to answer this question.

# Xinhua’s Global Horizons

*Xinhua* is firmly under the control of the CCP’s Central Propaganda Department. In 2013, President Xi Jinping exhorted the Propaganda Department to “tell Chinese stories well,” and the *Xinhua* news agency is at the forefront [@xinhua_xi_2013].  It also has an increasingly global reach to produce millions of articles in the past decade with hundreds of bureaus worldwide. Yet, *Xinhua*’s news reports should not be dismissed as mere propaganda or empty slogans. *Xinhua* presents itself as a balanced and fact-based news agency, not a government mouthpiece [@brazys_chinas_2020]. This outward-facing Chinese official media, whose mission as a key disseminator of the Party’s narrative to foreign audiences, may inform us of the PRC government’s initiatives of presenting Chinese perspectives of real-time events to global audiences.

*Xinhua* is global, but also serves as the main domestic media outlet. International and domestic audiences are two separate audiences, as *Xinhua* will produce its news contents in English, French, Spanish, Korean and Japanese, as well as Chinese. Brazys and Dukalaskis [-@brazys_chinas_2020] analyze how different countries are talked about in *Xinhua*’s articles in Chinese and in languages other than Chinese. They find that domestic-facing articles about other countries or regions are neither negative nor neutral, while articles about China are slightly positive. In contrast, *Xinhua*’s outward-facing articles use overall positive tones to report other countries and regions, including the United States. When characterizing China to foreign audiences, the tone is much more positive than the tone it uses for Chinese readers, or stories about any other country to both audiences.

The Chinese government is acutely sensitive to China’s image in foreign media coverage. In addition to building *Xinhua* as a global news agency, China also invests in ownership stake in media companies in other developing countries. A working paper by Sundquist [-@sundquist_telling_2022] shows that the media outlets after the ownership change adopt a more friendly tone when reporting China. This finding is consistent with the conventional argument that media ownership exerts a homogenizing effect on the newspapers’ coverage of foreign policy [@baum_media_2019]. Sending constructive media messages and shaping stories about China to international audiences is critical to Beijing, and *Xinhua* is not alone in this initiative.

>>>>

Although Beijing has primarily focused on its immediate surrounding neighbors, Asia-Pacific and Central Asia, for security and geostrategic concerns, it is nevertheless increasingly thinking and acting globally [@nathan_chinas_2012]. Similarly, in terms of where China's money goes and its economic influence is concentrated, some countries clearly are more important than others. However, Beijing's leaders hold the reasons why close to the vest.

@@

However, the goals of facilitating cooperation and winning international recognition as a rising power are difficult to accomplish in secret. This, therefore, raises a natural question. *Do China's foreign aid and debt allocation relate to public diplomacy and international recognition efforts? If so, how?*

@@

As we test these hypotheses, in our research design we take care to ensure that we delineate China's development financing according to the type of assistance China offers. Until recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) have unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.[^1] Thus, as we test our hypotheses, we also distinguish between Beijing-as-donor from Beijing-as-lender.

[^1]: See @dreher_banking_2022.

@@

Aid recipient countries disproportionately get more media attention from *Xinhua* via stories that report both on the state of their economy and their development needs. Notably, *Xinhua* does not just promote China's efforts to offer aid to those countries in need but also gives a rounded picture of those countries' economic conditions justifying their legitimate need for aid.

Meanwhile, the finding that loans correspond with diplomatic missions has special relevance in light of Beijing's Belt and Road Initiative (BRI). A 2018 BRI foreign policy assessment report from Washington [@hurley_examining_2019] highlights that lending indicates Beijing's economic interests and strategic dominance---China tends to lend to vulnerable states with debt distress in exchange for unfair deals and strategic concessions. Brautigam [-@brautigam_critical_2020] calls the Chinese "debt-trap diplomacy" a meme based on collective fear of the Chinese threat rather than sound empirical evidence. Our finding suggests that China is more likely to provide loans to diplomatically aligned countries. For instance, Russia, one of China's closest allies, receives the highest amount of lending on average among 142 countries but no aid at all between 2002 and 2017. In contrast, Iraq shows the opposite pattern: it receives the highest amount of aid from China during the period of our investigation, but it does not receive any loans. Concern about Chinese debt nonetheless raises a natural question related to the function of diplomacy in debtor countries. Does debt serve as a tool of diplomacy, or does Beijing wield diplomacy to cover a multitude of "debt-trap" sins? More research is needed to answer this question.

# Motivations behind Chinese Foreign Aid

China's motivations for giving foreign aid to developing countries mirror, in many ways, the motivations of Western donors. While recipient needs, in part, predicate Chinese assistance, so do Beijing's political and economic interests. Evidence suggests that China targets more finance toward poorer countries and does not systematically favor authoritarian over democratic recipients [@dreherEtAl2018]. At the same time, China gives more aid to countries that import more goods from China, while it gives less aid to those that officially recognize Taiwan [@dreherFuchs2015]. Further, China's Official Financing disproportionately goes to countries with more natural resources, greater UN General Assembly voting alignment with Beijing, and with more capacity to repay loans [@dreherEtAl2021].

Two waves of surveys of policymakers in 126 countries demonstrate the diplomatic impact that these investments have had.[^2] Compared to a first survey wave in 2014, responses in 2017 showed an increase in evaluations of China's influence with policymakers in developing countries. It would be naive to suppose this is a side-effect, rather than a goal, of China's development finance and other investments.

[^2]: Reported by AidData ["China's Financial Statecraft: Winning Africa one Yuan at a Time?"](https://www.aiddata.org/blog/chinas-financial-statecraft-winning-africa-one-yuan-at-a-time) on March 22, 2018. Accessed March 3, 2022

If its diplomatic motives for giving aid parallel those of other donor governments, visibility is surely a major concern for Beijing. Visual branding of aid has emerged as a growing concern among donors. The United States Agency for International Development (USAID), for example, cited its "new 'brand identity'" as instrumental in promoting the visibility of 2004-2005 tsunami relief to Indonesia.[^3] A US State Department report from 2004 further claims that better branding contributed to a doubling of favorable attitudes toward the US. Pew Research supports this view, finding that nearly 80$\%$ of Indonesians surveyed "said that post-tsunami aid from the US had improved their impression of America."[^4]

[^3]: See USAID's resource page on branding: [usaid.gov/branding](https://www.usaid.gov/branding).

[^4]: ["Does humanitarian aid improve America's image?"](https://www.pewresearch.org/global/2012/03/06/does-humanitarian-aid-improve-americas-image/) Published by Pew Research March 6, 2012. Accessed March 3, 2022.

Given the diplomatic uses of foreign aid, the obscurity that surrounds Beijing's development finance is all the more puzzling. Unlike Western donors, China does not report its financing activities to an international organization or make its activities visible in an online database. To say transparency is out of character for Beijing is an understatement, but foreign aid is one issue area where transparency would arguably serve China's interests. So why the secrecy?

As @fuchsRudyak2019 point out, the obscurity of China's development finance creates the perception among Western donors that China purposefully keeps its aid giving a secret. But as several scholars note, lack of transparency may have more to do with capacity and logistics than intention.[^5] For many decades, a complex and fractured bureaucracy has orchestrated Beijing's development financing. Only in the past few years, as its goals have become ever loftier, did China establish its first true bilateral aid agency---the China International Development Cooperation Agency (CIDCA). However, while established in 2018, much work remains to consolidate the operation and management of China's bilateral aid.[^6]

[^5]: See @fuchsRudyak2019 for several related citations.

[^6]: See this excellent summary entitled ["The Logic Behind China's Foreign Aid Agency"](https://carnegieendowment.org/2019/05/21/logic-behind-china-s-foreign-aid-agency-pub-79154) by the Carnegie Endowment for International Peace published May 21, 2019. Accessed March 10, 2022.

Absent transparency in reporting, there are other tools at Beijing's disposal that may complement the diplomatic objectives behind its foreign aid. In this study, we explore the connections between China's bilateral development finance and other, *visible*, activities linked to diplomatic and legitimacy-seeking objectives. We focus in particular on two factors: (1) diplomatic visits from China to developing countries and (2) coverage of developing countries in external-facing media (that is, media coverage intended for foreign, rather than domestic, audiences). We summarize these factors in greater detail in the following sections.

# Xinhua's Global Horizons

*Xinhua* is firmly under the control of the CCP's Central Propaganda Department. In 2013, President Xi Jinping exhorted the Propaganda Department to "tell Chinese stories well," and the *Xinhua* news agency is at the forefront [@xinhua_xi_2013]. It also has an increasingly global reach to produce millions of articles in the past decade with hundreds of bureaus worldwide. Yet, *Xinhua*'s news reports should not be dismissed as mere propaganda or empty slogans. *Xinhua* presents itself as a balanced and fact-based news agency, not a government mouthpiece [@brazys_chinas_2020]. This outward-facing Chinese official media, whose mission as a key disseminator of the Party's narrative to foreign audiences, may inform us of the PRC government's initiatives of presenting Chinese perspectives of real-time events to global audiences.

*Xinhua* is global, but also serves as the main domestic media outlet. International and domestic audiences are two separate audiences, as *Xinhua* will produce its news contents in English, French, Spanish, Korean and Japanese, as well as Chinese. Brazys and Dukalaskis [-@brazys_chinas_2020] analyze how different countries are talked about in *Xinhua*'s articles in Chinese and in languages other than Chinese. They find that domestic-facing articles about other countries or regions are neither negative nor neutral, while articles about China are slightly positive. In contrast, *Xinhua*'s outward-facing articles use overall positive tones to report other countries and regions, including the United States. When characterizing China to foreign audiences, the tone is much more positive than the tone it uses for Chinese readers, or stories about any other country to both audiences.

The Chinese government is acutely sensitive to China's image in foreign media coverage. In addition to building *Xinhua* as a global news agency, China also invests in ownership stake in media companies in other developing countries. A working paper by Sundquist [-@sundquist_telling_2022] shows that the media outlets after the ownership change adopt a more friendly tone when reporting China. This finding is consistent with the conventional argument that media ownership exerts a homogenizing effect on the newspapers' coverage of foreign policy [@baum_media_2019]. Sending constructive media messages and shaping stories about China to international audiences is critical to Beijing, and *Xinhua* is not alone in this initiative.
